### 2019-05-01 src/filter.tex
added src dir

\section{Filter design and implementation}\label{sec:filter}

\subsection{Block-parametric design}\label{sec:aaaa}

\subsection{Julia implementation}\label{sec:aaaa}

\subsection{Code optimization}\label{sec:aaaa}

\subsection{Performance analysis}\label{sec:aaaa}