### 2019-05-01 src/main.tex

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

%SetFonts
\usepackage[T1]{fontenc}
\usepackage{lmodern}

%SetFonts

\title{Fast algebraic filtering of surfaces from 3D medical images with Julia}
\author{Miroslav Ji\v{r}\'ik and Alberto Paoluzzi}
\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle

\begin{abstract}
In this paper we introduce a novel algebraic \textsc{lar-surf} filter, well founded on algebraic topology methods, to extract and smooth the boundary surface of any subset of voxels arising from the segmentation of a 3D medical image. The input is defined as a \emph{chain}, i.e.~as a vector from a linear space of 3-chains, represented in coordinates as a sparse Boolean vector. The output is produced as the result of the mapping via the linear boundary operator $\partial_3:C_3 \to C_2$ between linear spaces of 3- and 2-chains.  In particular, when the input set of voxels is either not (4-)connected, or contains one or more empty regions inside, \textsc{lar-surf} generates a non connected set of closed surfaces, i.e.~a set of 2-cycles---using the language of algebraic topology. The only data structures used by this approach are sparse arrays with one or two indices, i.e.~sparse vectors and matrices. This work is based on LAR (Linear Algebraic Representation) methods, and is implemented in Julia language, natively supporting parallel computing on hybrid architectures.
\end{abstract}

\tableofcontents

%\section{}
%\subsection{}

\include{intro}
\include{background}
\include{filter}
\include{examples}
\include{conclusion}

\end{document}