### 2019-05-09 src/background.tex
working on algorithm design

\section{Background}\label{sec:background}

\subsection{Linear Algebraic Representation}\label{sec:aaaa}

A \emph{representation scheme} for solid modeling is a mapping between a space of mathematical models and a space of symbolic representations, often generated by a formal grammar.
Solid pointsets (i.e., `$r$-sets') are defined~\cite{requicha:80} as compact (bounded and closed) regular and semianalytic subsets of the $d$-space. A large number of representation schemes were defined in the past forty years, including the two main classes of (a) \emph{boundary representations} (`$B$-reps'), where the solid model is represented through a representation of its boundary elements, i.e.~faces, edges and vertices, and (b) \emph{decompositive/enumerative representations}, that are a decomposition of either the object or the embedding space, respectively, into a well-defined cellular complex. In particular, a boundary representation provides a cellular decomposition of the object's boundary into cells of dimension zero (vertices), one (edges), and two (faces). Medical imaging is a kind of enumerative representation of organs and tissues of interest, as subsets of 3D volume elements (voxels) from the 3D image.

\subsection{Multiindices from Cartesian indices}\label{sec:aaaa}

\subsection{Taubin Smoothing}\label{sec:aaaa}