### 2019-05-10 src/implementation.tex
splitted filter.tex file into design.tex and implementation.tex

\subsection{Julia implementation}\label{sec:implementation}

\subsubsection{Code optimization}\label{sec:optimization}

\subsubsection{Performance analysis}\label{sec:analysis}