### 2019-07-07 src/examples.tex
new examples added and text with statistic

\section{Examples}\label{sec:examples}

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{figs/porta_smoothing_1.png} 
\vspace{0.05\textwidth}
\includegraphics[width=0.4\textwidth]{figs/porta_smoothing_2.png} 
%\includegraphics[scale=1]{input/ircad_comparison.pdf} 
\caption{Triangulated isosurface of portal vein calculated with LAR-SURF}
\label{fig:example_porta}
\end{figure}