### 2019-07-08 src/examples.tex

\section{Examples}\label{sec:examples}

The 
For an experiments the we used the dataset Ircad1b.

\begin{table}
\input{input/dataset_ircad}
\end{table}
\begin{table}
\input{input/dataset_ircad_desc}
\end{table}

We performed experiment on Ircad dataset to compare time requirements.

\begin{figure}
\includegraphics[width=0.99\textwidth]{figs/ircad_comparison.pdf} 
%\includegraphics[scale=1]{input/ircad_comparison.pdf} 
\caption{Time requirements of LAR-SURF filter and Marching Cubes on Ircad dataset. Error bars shows the 95\% confidence interval}
\end{figure}