### 2020-04-25 src/paper.tex

\documentclass[9pt,academicons]{article}

\usepackage{CADA}

%%%%%%%%%%%%%%%% original head %%%%%%%%%%%%%%%%%%

% \documentclass[11pt, oneside]{amsart}   	% use "amsart" instead of "article" for AMSLaTeX format
% \usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   		% ... or a4paper or a5paper or ...
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{hyperref}

%SetFonts
% \usepackage[utf8x]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage{lmodern}

\nonstopmode
\input{macros}

% rules for for inputed tables
% the
\usepackage{siunitx}
\usepackage{booktabs} % tables

%alternative rules for tables
%\newcommand{\toprule}{\hline}
%\newcommand{\midrule}{\hline}
%\newcommand{\bottomrule}{\hline}
%\newcommand{\specialrule}{}

%%%%%%% M. Jirik added - code highlithing
% \usepackage{minted}
% \usepackage[cache=false]{minted}
\usepackage{listings}
\usepackage[dvipsnames]{xcolor}
% \usepackage{minted}
%SetFonts
\lstdefinelanguage{Julia}%
  {morekeywords={abstract,break,case,catch,const,continue,do,else,elseif,%
      end,export,false,for,function,immutable,import,importall,if,in,%
      macro,module,otherwise,quote,return,switch,true,try,type,typealias,%
      using,while},%
   sensitive=true,%
%   alsoother={$},%
   morecomment=[l]\#,%
   morecomment=[n]{\#=}{=\#},%
   morestring=[s]{"}{"},%
   morestring=[m]{'}{'},%
}[keywords,comments,strings]%

\lstset{%
    language         = Julia,
    basicstyle       = \ttfamily,
    keywordstyle     = \bfseries\color{blue},
    stringstyle      = \color{magenta},
    commentstyle     = \color{ForestGreen},
    % commentstyle     = \color{green},
    showstringspaces = false,
}
%%%%%%
% TODO remove
\usepackage{todonotes}

\title{Algebraic Filtering of Surfaces from 3D Medical Images with Julia}
% \author{Miroslav Jirik and Alberto Paoluzzi}
\date{}							% Activate to display a given date or no date

% \usepackage{draftwatermark}
% \SetWatermarkText{Draft}
% \SetWatermarkScale{2.0}
% \SetWatermarkColor[rgb]{0.9,0.9,0.9}

\begin{document}
\maketitle

% ############################################
% Here is the fun part. Please enter your:            
%      Name (First, Middle initial, Last)                  
%      Your ORCID number replace the "000-0000-1234-5678" part.    
% ############################################

\authorSection{
	\anAuthor{Miroslav Jirik}{0000-0002-8002-2079}{1,3},
	\anAuthor{Alberto Paoluzzi}{0000-0002-3958-8089}{2}, 
% 	\anAuthor{First M. Last}{000-0000-1234-5678}{3} 
}

% ######################################
% Here we need your affiliation and contact e-mail. Please edit   
% affiliation as well as the e-mail fields.                                  
% ######################################
\affiliationSection{
	\anAffiliation{1}{
	Charles University,
    Faculty of Medicine in Pilsen, 
    Biomedical Center}{jirik@lfp.cuni.cz}
	\anAffiliation{2}{Roma Tre University, Department of Mathematics and Physics}{paoluzzi@dia.uniroma3.it}
	\anAffiliation{3}{NTIS, Faculty of Applied Sciences, University of West Bohemia}{}
}

% ######################################
% Please decide on who the corresponding author is going to be 
% and complete the section below                                            
% ######################################
\correspondingAuthor{Miroslav Jirik}{jirik@lfp.cuni.cz}

% ########################################
% Please type in your abstract below after the "\abstract{" part.
% ########################################

\abstract{
In this paper we introduce a novel algebraic 
% \textsc{lar-surf} 
filter, well-founded on algebraic topology methods, to extract and smooth the boundary surface of any subset of voxels arising from  the segmentation of a 3D medical image. 
% TODO 3D medical image
The input of our Linear Algebraic Representation Surface extraction filter (\textsc{lar-surf}) is defined as a \emph{chain}, i.e.~as a vector from a linear space of 3-chains, represented in coordinates as a sparse binary vector. The output is produced by a linear mapping between spaces of 3- and 2-chains through the boundary operator $\partial_3:C_3 \to C_2$.  
% In particular, when the input set of voxels is either not (4-)connected, or contains one or more empty regions inside, \textsc{lar-surf} generates a non-connected set of closed surfaces, i.e. quadrilateral faces or~a set of 2-cycles---using the language of algebraic topology. 
The only data structures used by this approach are sparse arrays with one or two indices, i.e.~sparse vectors and matrices. This work is based on Linear Algebraic Representation methods and is implemented in Julia language, natively supporting parallel  computing on hybrid hardware architectures.
}

\keywords{
Medical 3D, Computational Topology, Linear Algebraic Representation, LAR, Julia, Surface Extraction
}

% DOI: 10.14733/cadconfP.2020.xxx-yyy
\doi{10.3722/cadaps.2020.xxx-yyy}

% \tableofcontents

\input{intro}
\input{background}
\input{design}
\input{implementation}
\input{examples}
\input{conclusion}

\section*{ACKNOWLEDGEMENTS}
\sloppy{
This work was supported by Charles University Research Centre program UNCE/MED/006 ``University Center of Clinical and Experimental Liver Surgery'' 
and Ministry of Education project ITI CZ.02.1.01/0.0/0.0/17\_048/0007280: Application of modern technologies in medicine and industry.
} 
The research was also supported by the project LO 1506 of the Czech Ministry of Education, Youth and Sports.
% The authors appreciate the access to computing and storage facilities owned by parties and projects contributing to the National Grid Infrastructure MetaCentrum  provided under the program ``Projects of Large Research, Development, and Innovations Infrastructures'' (CESNET LM2015042).

\appendix
\input{appendix}

\section*{ORCID}
\orcid{Miroslav Jirik}{0000-0002-8002-2079}
\orcid{Alberto Paoluzzi}{0000-0002-3958-8089}

% \clearpage{}
\referenceSection
\bibliographystyle{CADA}
% \bibliography{library,TSAS-2019,references}
\bibliography{TSAS-2019,references}

\bigskip
\end{document}