### 2020-04-25 src/examples.tex

\section{Examples}\label{sec:examples}

% \includegraphics[height=0.3\textwidth]{figs/nrn10_100_green.png} 
%\includegraphics[scale=1]{input/ircad_comparison.pdf} 
\begin{figure}
\centering
% \includegraphics[width=0.3\textwidth]{figs/ircad01}
\includegraphics[width=0.4\textwidth]{figs/liver_01_red_3.png} 
% \vspace{0.05\textwidth}
\includegraphics[width=0.4\textwidth]{figs/portalvein_01_yellow_3.png} 
% \vspace{0.01\textwidth}
\includegraphics[height=0.4\textwidth]{src/figs/nrn10_100_magenta_high_res.png} 
% \includegraphics[height=0.4\textwidth]{figs/nrn10_100_low_res.png} 
% \vspace{0.05\textwidth}
\includegraphics[width=0.4\textwidth]{figs/porta_smoothing_2.png} 
\caption{
Triangulated isosurface of liver parts extracted with LAR-SURF algorithm. 
On the upper left image the surface of the liver parenchyma of human can be seen. 
Both right images show the Portal Vein from the same Ircad dataset \cite{ircad} 
with resolution $1.6\times0.57\times0.57$ [mm].
On the left bottom image is the pig liver microvasculature based on corrosion cast prepared by Eberlova
\cite{eberlova2017use}. The resolution of the Micro-CT data is 4.682 $\mu{}m$.
} \label{fig:example_liver_macro_micro}
\end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=0.4\textwidth]{figs/porta_smoothing_1.png} 
% \vspace{0.05\textwidth}
% \includegraphics[width=0.4\textwidth]{figs/porta_smoothing_2.png} 
% %\includegraphics[scale=1]{input/ircad_comparison.pdf} 
% \caption{Triangulated isosurface of portal vein calculated with LAR-SURF}
% \label{fig:example_porta}
% \end{figure}

The implementation of our algorithm is available in LarSurf package. 
Use of this package can be seen on listings \ref{lst:example1} where the liver segmentation with 2865131 voxels from
Ircad dataset is used as an input for our surface extraction algorithm. The size of 3D volumetric 
image is $129 \times 512 \times 512$
and the voxel resolution is $1.6\times0.57\times0.57$ [mm]. 
The output surface model created by  182124 triangles and the number of vertices is 90822. The visualization can 
be seen on figure \ref{fig:example_liver_macro_micro}. 
The portal vein surface extraction can be performed with a small change of input path in code.
The The 3D image resolution is the same. 
It can be seen in the same figure. 
The number of input voxels is 103533. 
The output surface is created by 90822 vertices and 182124 triangles.

\begin{lstlisting}[caption={Get surface from DICOM volumetric data}, label={lst:example1}]
using Distributed
using Pio3d  # Read 3D data from DICOM files
addprocs(3)  # set number of processors
using LarSurf

LarSurf.lsp_setup([64, 64, 64])  # set block size

# read data from DICOM files
datap = Pio3d.read3d("3Dircadb1.1/MASKS_DICOM/liver")
segmentation = datap["data3d"]
voxelsize_mm = datap["voxelsize_mm"]

# get surface
V, FV = LarSurf.lsp_get_surface(segmentation, voxelsize_mm)
FVtri = LarSurf.triangulate_quads(FV)

# do smoothing and save data
Vs = LarSurf.Smoothing.smoothing_FV_taubin(V,FV,0.5,-0.2,40)
objlines = LarSurf.Lar.lar2obj(Vs, FVtri, "liver.obj")
\end{lstlisting}

The right image of figure \ref{fig:example_liver_macro_micro} is surface of a microvasculature of pig liver. 
Volumetric image is based on Micro-CT data of corrosion casts of pig liver.
\cite{eberlova2017use}.
The size of the visualized data is $100\times100\times100$ voxels and size of voxel is 4.682 $\mu{}m$.
The number of of triangles is 
544784 and number of vertices is 272826.

% \begin{minted}{python}

% using Pkg
% \end{minted}
% \begin{minted}[breaklines,escapeinside=||,mathescape=true, linenos, numbersep=3pt, gobble=2, frame=lines, fontsize=\small, framesep=2mm]{julia}
% Your awesome julia code here\end{minted}